---
title: "AMS380 Project"
author:
  - "Alex Takamizawa"
  - "Ethan Hoskins"
  - "Jiarui Zha"
  - "Tianrun Gou"
format: html
editor: visual
---

## Environment Setup

### Loading Packages

```{r}
#| output: false
install.packages(
  c("AER", "caret", "DescTools", "hydroTSM", "MASS", "nnet", "tidyverse")
)

library(AER)
library(caret)
library(DescTools)
library(hydroTSM)
library(MASS)
library(nnet)
library(tidyverse)
```

### Setting Seed

```{r}
set.seed(123)
```

### Loading and Preparing Data

```{r}
bird_collisions <- read.csv("data/bird_collisions.csv") %>%
  select(-c("genus", "species")) %>%
  subset(
    !family %in% c(
      "Icteridae",
      "Icteriidae",
      "Laniidae", 
      "Polioptilidae",
      "Sittidae",
      "Tyrannidae",
      "Vireonidae"
    )
  )
bird_collisions$date <- time2season(
  as.Date(bird_collisions$date),
  out.fmt = "seasons"
)
```

## Multinomial Logistic Regression

### Partitioning Data

```{r}
train_idx <- bird_collisions$family %>%
  createDataPartition(p = 0.70, list = FALSE)
train <- bird_collisions[train_idx, ]
test <- bird_collisions[-train_idx, ]
```

### Fitting the Model

```{r}
model_multinom <- multinom(family ~ ., data = train)

summary(model_multinom)
```

### Evaluating the Model

```{r}
pred_classes <- model_multinom %>% predict(test)

mean(pred_classes == test$family)
```

## Generating Count Data

### Generating Data

```{r}
counts <- bird_collisions %>%
  count(date, locality, family, flight_call, habitat, stratum, sort = TRUE)
```

### Preparing Count Data

```{r}
# counts$date <- as.factor(counts$date)
# counts$locality <- as.factor(counts$locality)
# counts$family <- as.factor(counts$family)
# counts$flight_call <- as.factor(counts$flight_call)
# counts$habitat <- as.factor(counts$habitat)
# counts$stratum <- as.factor(counts$stratum)
# counts$n <- as.numeric(counts$n)
```

### Visualizing Count Data

```{r}
ggplot(counts, aes(family, n)) + geom_point(aes(colour = family))
```

```{r}
ggplot(counts, aes(x = family)) + geom_bar()
```

```{r}
ggplot(counts, aes(n, fill = habitat)) +
  geom_histogram() +
  scale_x_log10() +
  facet_grid(habitat ~ ., margins = TRUE, scales = "free_y")
```

## Poisson Regression

### Fitting the Model

```{r}
model_pois <- glm(n ~ ., data = counts, family = poisson)

summary(model_pois)
```

### Dispersion Test

```{r}
dispersiontest(model_pois)
```

The p-value resulting from running an overdispersion test on the Poisson model is less than the significance level (0.05), so we reject the null hypothesis. Therefore, the true dispersion of the count data is greater than 1. As the count data is overdispersed, Quasi-Poisson regression or Negative Binomial regression would yield a more effective model.

```{r}
mean(counts$n)
var(counts$n)
```

The variance of the count data is significantly greater than the mean, therefore a Negative Binomial model would work best given the count data.

## Negative Binomial Regression

### Fitting the Model

```{r}
model_neg_binom <- glm.nb(n ~ ., data = counts)

summary(model_neg_binom)
```

### Plotting Residuals

```{r}
plot(
  model_neg_binom$fitted.values,
  model_neg_binom$residuals,
  pch = 20,
  xlab = "Collision Count",
  ylab = "Standatdized Residuals",
  main = "Negative Binomial"
)
abline(0, 0)
```

### Evaluating Negative Binomial Model

```{r}
PseudoR2(model_neg_binom)
```
